---
title: "SEM_paper"
output: pdf_document
---

```{r setup, include=FALSE}
# importing packages
library("haven")        
library("dplyr")      
library("psych")
library('stringr')
library("lavaan")   
library("tidySEM")
library("purrr") 
library(ggplot2)
library("corrplot")
library("semTools")
library(MVN)
library(Amelia)
```

# data importing
```{r}
data_2017=read.csv('../ESS7e02_2.csv')

belgium_2017 <- data_2017 %>% filter(cntry=="BE") %>% select(imbgeco, imueclt, imwbcnt, imtcjob, imbleco,imwbcrm,qfimedu, qfimlng, qfimchr,qfimwht, qfimwsk,qfimcmt,stflife, stfeco, stfgov, stfedu, stfdem,agea, gndr, eduyrs, hinctnta)


write.csv(belgium_2017, "ESS7_belgium2017.csv")
```

# description of data
```{r}
ess=read.csv("ESS7_belgium2017.csv", row.names = 1)
es.std=scale(ess)

descriptive_ess= as.data.frame(describe(ess))
descriptive_ess
```

# Correlation and covariance table of variables
```{r}
# correlation graph for welfare support government responsibility variables
ess.imm=ess%>%select(imbgeco, imueclt, imwbcnt, imtcjob, imbleco,imwbcrm)
imm.cov<- cov(ess.imm, use = "pairwise.complete.obs")
imm.cor <- cov2cor(imm.cov)
corrplot::corrplot(imm.cor, 
                   is.corr = FALSE, 
                   method = "circle", 
                   type = "upper" )

# correlation graph for welfare support criticism variables
ess.condition=ess%>%select(qfimedu, qfimlng, qfimchr,qfimwht, qfimwsk,qfimcmt)
condition.cov<- cov(ess.condition, use = "pairwise.complete.obs")
condition.cor <- cov2cor(condition.cov)
corrplot::corrplot(condition.cor, 
                   is.corr = FALSE, 
                   method = "circle", 
                   type = "upper" )

# correlation graph for satisfaction variables
ess_df.satif=ess_df%>%select(stflife, stfeco, stfgov, stfedu, stfhlth)
welfare.satif.cov<- cov(ess_df.satif, use = "pairwise.complete.obs")
welfare.satif.cor <- cov2cor(welfare.satif.cov)
corrplot::corrplot(welfare.satif.cor, 
                   is.corr = FALSE, 
                   method = "circle", 
                   type = "upper" )

# overall correlation graph
cov <- cov(ess, use = "pairwise.complete.obs")
cor <- cov2cor(cov)
corrplot::corrplot(cor, 
                   is.corr = FALSE,
                   method = "circle", 
                   type = "upper"   )

```

# detecting normality of data
1. using KS(Kolmogorov_Smirnov) test, if D is more close to 0 then the data is more normally distributed. 

2. using Henze-Zirkler's multivariate normality test, if p-value < 0.05 then non-normality
```{r}
#checking if the univariate is normal, if the univariate distribution is non-normality then multivariate distribution is not normal.

# qualification of immigration histogram
# not normally distributed
edu=ggplot(ess, aes(qfimedu))+geom_histogram(aes(y=..density..), binwidth=1, color='black', alpha=0.3)

work=ggplot(ess, aes(qfimwsk))+geom_histogram(aes(y=..density..), binwidth=1, color='black', alpha=0.3)

# social satisfaction scores histogram
life=ggplot(ess, aes(stflife))+geom_histogram(aes(y=..density..), binwidth=1, color='black', alpha=0.3)

econ=ggplot(ess, aes(stfeco))+geom_histogram(aes(y=..density..), binwidth=1, color='black', alpha=0.3)


#KS test to check the normality
ks_life=ks.test(ess$stflife, 'pnorm', mean=mean(ess$stflife, na.rm=T), sd=sd(ess$stflife, na.rm=T))
ks_life
# D=0.21458, p-value close to zero

ks_edu=ks.test(ess$qfimedu, 'pnorm', mean=mean(ess$qfimedu, na.rm=T), sd=sd(ess$qfimedu, na.rm=T))
ks_edu
# D=0.21708, p-value close to zero

ks_econ=ks.test(ess$imbgeco, 'pnorm', mean=mean(ess$imbgeco, na.rm=T), sd=sd(ess$imbgeco, na.rm=T))
ks_econ
# D=0.29409, p-value close to zero

data.frame(Variables= c("life_satisfaction", "education_qualification", "social_economy"),
          D = round(c(ks_life$statistic, ks_edu$statistic, ks_econ$statistic),2),
          "P-value" = c(ks_life$p.value, ks_edu$p.value, ks_econ$p.value))

# HZ test
ess_ws <- ess[,c("stflife", "qfimedu", "imbgeco")]
ess_ws_na <- na.omit(ess_ws)
mvn <- mvn(data = ess_ws_na,mvnTest = c("hz"))
mvn$multivariateNormality
# the result shows a non-multivariate normality, with HZ value=130.285, p-value close to zero. 

# checking missing values
ess_wc <- ess%>% select(imbgeco, imueclt, imwbcnt, imtcjob, imbleco,imwbcrm,qfimedu, qfimlng, qfimchr,qfimwht, qfimwsk,qfimcmt,stflife, stfeco, stfgov, stfedu, stfdem)
# Amelia explicitly requires a object of data.frame class 
ess_wc <- data.frame(ess_wc)

a.out <- amelia(ess_wc, m = 15,seed = 23 )
# the output result shows there is no missing value in the dataset. 
```



# other CFA model (exercise 3)
1. test equal loadings (metirc invariance)
2. test eqaal loadings and intercepts (scalar invariance)
3. test equal loadings, intercepts, and residuals (strict invariance)
4. test factor variance and covariance (structural invariance)
5. evaluating measurement invariance, check which model is better
6. then checking modification indices, and free one parameter at a time "parTable(fit_strict)" can check which parameter need to be freed.
7. then run cfa again to compare the results


# model fit stats
1. CFI, TLI, RMSEA, SRMR
2. checking modification indices

# CFA configural model 

```{r}
model1  <-'
    # political impact of immigration
    impact=~imbgeco+imueclt+imwbcnt+imtcjob+imbleco+imwbcrm
    # immigration qualification
    eligibility=~ qfimedu+qfimlng+qfimchr+qfimwht+qfimwsk+qfimcmt
    #social _satisfaction
    satisfaction =~ stflife+ stfeco+ stfgov+ stfedu+ stfdem

    impact~~1*impact
    eligibility~~1*eligibility
    satisfaction~~1*satisfaction
    impact~~eligibility
    impact~~satisfaction
    eligibility~~satisfaction
'
fit.model=cfa(model1, data = ess, estimator="ML")
fit.model.rob<- cfa(model1, data = ess, estimator="MLM")
summary(fit.model.rob,  fit.measures=T)
summary(fit.model, fit.measures=T)
#fitMeasures(fit.model.rob, c('chisq','df', 'pvalue', 'cfi',  'tli', 'rmsea', 'srmr'))
#fitMeasures(fit.model, c('chisq','df', 'pvalue', 'cfi',  'tli', 'rmsea', 'srmr'))

#plot the sem
lay <- get_layout("impact", "", "", "","","",
                  "eligibility","", "","", "","",
                  "satisfaction", "","", "", "",
                  "imbgeco","imueclt","imwbcnt","imtcjob","imbleco", "imwbcrm",
 "qfimedu", "qfimlng", "qfimchr", "qfimwht","qfimwsk","qfimcmt",
 "stflife", "stfeco", "stfgov", "stfedu","stfdem",rows = 2)

plot_wc <- graph_sem(model = fit.model.rob,   
                    layout = lay,  
                    angle = 180,text_size = 1.5 )   

plot_wc
modificationIndices(fit.model.rob,standardized=TRUE)

#imbgeco	~~	stfdem	36.846
#imtcjob	~~	qfimlng	54.028
#qfimchr	~~	stflife	38.636

# chi-square test has p-value=0 which indicate the model is lack of fit 
# but the chi-square test will be impacted by the sample size,
# from the CFI, TLI in standard result, they are lower than 0.90 which means lack of fit. the CFI and TLI improved in the robust result, although it still an indicator of lack of fit.
# the RMSEA and SRMR in the robust model meet the criteria of goodness of fit. 
# in this case we have a mix evidence of evaluating the overall , and compare to the standard model, robust model is better.

# next we see the factor loading, the first factor loading is assigned to 1 to scale each latent variable. 
# then the p-value shows the significance of factor loading, we can see most of the variables are significant except qfimchr. we would consider about removing this variables from the model and check the goodness of fit. 
# std.all is the correlation value

# since eligibility~~satisfaction does not have significant relationship, we remove the relationship from the model to test the goodness of fit. 


```

# remove qfimchr from eligibility
```{r}
model2  <-'
    # political impact of immigration
    impact=~imbgeco+imueclt+imwbcnt+imtcjob+imbleco+imwbcrm
    # immigration qualification, remove qfimchr
    eligibility=~ qfimedu+qfimlng+qfimwht+qfimwsk+qfimcmt
    #social _satisfaction
    satisfaction =~ stflife+ stfeco+ stfgov+ stfedu+ stfdem

    impact~~1*impact
    eligibility~~1*eligibility
    satisfaction~~1*satisfaction
    impact~~eligibility
    impact~~satisfaction
    #eligibility~~satisfaction
'
#imbgeco	~~	stfdem	36.846
#imtcjob	~~	qfimlng	54.028
#qfimchr	~~	stflife	38.636
fit.model2<- cfa(model2, data = ess, estimator='MLM')
summary(fit.model2, standardized=TRUE, fit.measures=T)
fitMeasures(fit.model2, c('chisq','df', 'pvalue', 'cfi',  'tli', 'rmsea', 'srmr'))
modificationIndices(fit.model2,standardized=TRUE)
lay <- get_layout("impact", "", "", "","","",
                  "eligibility","", "","", "","",
                  "satisfaction", "","", "", "",
                  "imbgeco","imueclt","imwbcnt","imtcjob","imbleco", "imwbcrm",
 "qfimedu", "qfimlng", "qfimchr", "qfimwht","qfimwsk","qfimcmt",
 "stflife", "stfeco", "stfgov", "stfedu","stfdem",rows = 2)

plot_m2 <- graph_sem(model = fit.model2,   
                    layout = lay,  
                    angle = 180,text_size = 1.5 ) 
plot_m2

 
```

```{r}
# model 3, base on the modification measurements, add some relationship between predictors
model3  <-'
    # political impact of immigration
    impact=~imbgeco+imueclt+imwbcnt+imtcjob+imbleco+imwbcrm
    # immigration qualification, remove qfimchr
    eligibility=~ qfimedu+qfimlng+qfimwht+qfimwsk+qfimcmt
    #social _satisfaction
    satisfaction =~ stflife+ stfeco+ stfgov+ stfedu+ stfdem

    impact~~1*impact
    eligibility~~1*eligibility
    satisfaction~~1*satisfaction
    impact~~eligibility
    impact~~satisfaction
    eligibility~~satisfaction
    
    imtcjob	~~	qfimlng
'
fit.model3<- cfa(model3, data = ess, estimator='MLM')
summary(fit.model3, standardized=TRUE, fit.measures=T)

# in model 3 we add the relationship between imtcjob and qfimlng. from the covariances result, the significant p-value states the latent variables impact and eligibility has a negative relationship. However, the two predictors from the latent variables do not have any relationship, since the p-value=0.260 which greater than 0.05. Thus it is not necessary to add the relationship into the model. Also we can see that model3 does not improve a large amount compared to model2, then we will use model2 in further MIMC analysis.  


```

```{r}
# observed covaraince matrix
lavInspect(fit.model2, "sampstat")

# estimated covariance matrix
fitted(fit.model2)

# difference between the observed and estimated covariance matrix
residuals(fit.model2)

# standard errors for free parameters in the model
lavTech(fit.model2, "se")

# parameter estimates with confidence interval
parameterEstimates(fit.model2, standardized=TRUE)

```



# MIMC model (multiple indicators, multiple causes), Ex2
1. which variable in agea,eduyrs,gndr,hinctnta have effect on government and social variable. 

```{r}
# model.mimic is based on model2, without 2 variables in wc
# all additional effects on ws
model2.mimic  <-'
    # political impact of immigration
    impact=~imbgeco+imueclt+imwbcnt+imtcjob+imbleco+imwbcrm
    # immigration qualification, remove qfimchr
    eligibility=~ qfimedu+qfimlng+qfimwht+qfimwsk+qfimcmt
    #social _satisfaction
    satisfaction =~ stflife+ stfeco+ stfgov+ stfedu+ stfdem

    impact~~1*impact
    eligibility~~1*eligibility
    satisfaction~~1*satisfaction
    impact~~eligibility
    impact~~satisfaction
    eligibility~~satisfaction
    
    impact~ agea+eduyrs+gndr+hinctnta
    eligibility~ agea+eduyrs+gndr+hinctnta
    satisfaction~ agea+eduyrs+gndr+hinctnta
'

fit2.mimic=cfa(model2.mimic, data=ess, estimator='MLM')
summary(fit2.mimic, standardized=TRUE, fit.measures=T)
#fitMeasures(fit.mimic, c('chisq','df', 'pvalue', 'cfi',  'tli', 'rmsea', 'srmr'))

model2.mimic2  <-'
    # political impact of immigration
    impact=~imbgeco+imueclt+imwbcnt+imtcjob+imbleco+imwbcrm
    # immigration qualification, remove qfimchr
    eligibility=~ qfimedu+qfimlng+qfimwht+qfimwsk+qfimcmt
    #social _satisfaction
    satisfaction =~ stflife+ stfeco+ stfgov+ stfedu+ stfdem

    impact~~1*impact
    eligibility~~1*eligibility
    satisfaction~~1*satisfaction
    impact~~eligibility
    impact~~satisfaction
    eligibility~~satisfaction
    
    impact~ eduyrs
    eligibility~ agea+eduyrs
    satisfaction~ hinctnta
'

fit2.mimic2=cfa(model2.mimic2, data=ess, estimator='MLM')
summary(fit2.mimic2, standardized=TRUE, fit.measures=T)
# only education years will affect the impact of immigration
# age and education years will affect the qualification of immigration
# household income will affect the satisfaction of public

lay <- get_layout( "qfimedu", "qfimlng",  "qfimwht","qfimwsk","qfimcmt","","","",
                   "","agea","eligibility","","","eduyrs","","","",
                   "","","","", "impact","","","","",
                   "","","","","imbgeco","imueclt","imwbcnt","imtcjob","imbleco","imwbcrm",
                   "", "satisfaction","", "hinctnta","", "", "","", "", 
 "stflife", "stfeco", "stfgov", "stfedu","stfdem","","","","",rows = 6)
lay

graph_sem(model = fit2.mimic2,   
                    layout = lay,  
                    text_size = 1.5 ) 
```

# configural, metric, strong test
```{r}
cfa.config <- cfa(model2.mimic2, data = ess,estimator="MLM",  group="gndr")
summary(cfa.config, fit.measures = TRUE, standardized = TRUE)

# Metric model
cfa.metric <- cfa(model2.mimic2,data = ess, estimator="MLM",group = "gndr", group.equal = "loadings")
summary(cfa.metric, fit.measures = TRUE, standardized = TRUE)

# Model comparison
anova(cfa.config, cfa.metric)
#This finding suggests that after constraining the factor loadings to be equal across groups, the model fit did change substantially.

# Scalar model
cfa.scalar <- cfa(model2.mimic2, data = ess, group = "gndr",estimator="MLM", group.equal = c("loadings","intercepts"))
summary(cfa.scalar, fit.measures = TRUE, standardized = TRUE)
# Model comparison
anova(cfa.metric, cfa.scalar)
#This finding suggests that after constraining the factor loadings and intercept to be equal across groups, the model fit did not change substantially. 
# metirc model is better than other 2

rbind(config=fitMeasures(cfa.config, c('chisq','df', 'pvalue', 'cfi',  'tli', 'rmsea', 'srmr')),
metric=fitMeasures(cfa.metric, c('chisq','df', 'pvalue', 'cfi',  'tli', 'rmsea', 'srmr')),
scalar=fitMeasures(cfa.scalar, c('chisq','df', 'pvalue', 'cfi',  'tli', 'rmsea', 'srmr')))

# from the result we can see that the goodness of fit for 3 models deviate from the perfect model. Comparing the fit measurements, CFI and TLI of metric model and scalar model decrease as we put equal loading and intercept constrains. But the RMSEA and SRMR meet the fit criteria (above 0.95) after adding equal loading constrain, equal intercept condition does not make any changes to improve the model.  

# need to use other model to test
```

# mediation analysis Ex2
1. direct and indirect effect of variables

```{r}
# without socio features impact
model2.med='
    # political impact of immigration
    impact=~imbgeco+imueclt+imwbcnt+imtcjob+imbleco+imwbcrm
    # immigration qualification, remove qfimchr
    eligibility=~ qfimedu+qfimlng+qfimchr+qfimwht+qfimwsk+qfimcmt
    #social_satisfaction
    satisfaction=~stflife+ stfeco+ stfgov+ stfedu+ stfdem

    impact~~1*impact
    eligibility~~1*eligibility
    satisfaction~~1*satisfaction
    impact~~eligibility
    impact~~satisfaction

    # direct effect
    satisfaction ~ a*impact
    
    # mediator
    eligibility ~ b*impact
    satisfaction ~ c*eligibility

    #indirect effect
    indirect:=b*c
    
    # total effect
    total:=a+(b*c)
'
fit.med=cfa(model2.med, data=ess, estimator='MLM',orthogonal = TRUE)
summary(fit.med,standardized=TRUE, fit.measures=T)
lay <- get_layout( "qfimedu", "qfimlng",  "qfimwht","qfimwsk","qfimcmt","","","",
                   "","","eligibility","","","","","","",
                   "","","","", "impact","","","","",
                   "","","","","imbgeco","imueclt","imwbcnt","imtcjob","imbleco","imwbcrm",
                   "", "satisfaction","", "","", "", "","", "", 
 "stflife", "stfeco", "stfgov", "stfedu","stfdem","","","","",rows = 6)
lay
graph_sem(model = fit.med,   
                    layout = lay,  
                    text_size = 1.5 ) 


```


```{r}
model2.mimic.med  <-'
    # political impact of immigration
    impact=~imbgeco+imueclt+imwbcnt+imtcjob+imbleco+imwbcrm
    # immigration qualification, remove qfimchr
    eligibility=~ qfimedu+qfimlng+qfimwht+qfimwsk+qfimcmt
    #social _satisfaction
    satisfaction =~ stflife+ stfeco+ stfgov+ stfedu+ stfdem

    impact~~1*impact
    eligibility~~1*eligibility
    satisfaction~~1*satisfaction
    impact~~eligibility
    impact~~satisfaction
    
    impact~ eduyrs 
    eligibility~ agea+eduyrs
    satisfaction~ hinctnta
    
    # direct effect
    satisfaction ~ a*impact
    
    # mediator
    eligibility ~ b*impact
    satisfaction ~ c*eligibility

    #indirect effect
    indirect:=b*c
    
    # total effect
    total:=a+(b*c)
'
fit.mimic.med=cfa(model2.mimic.med, data=ess, estimator="MLM")
summary(fit.mimic.med,standardized=TRUE, fit.measures=T)
lay <- get_layout( "qfimedu", "qfimlng",  "qfimwht","qfimwsk","qfimcmt","","","",
                   "","agea","eligibility","","","eduyrs","","","",
                   "","","","", "impact","","","","",
                   "","","","","imbgeco","imueclt","imwbcnt","imtcjob","imbleco","imwbcrm",
                   "", "satisfaction","", "hinctnta","", "", "","", "", 
 "stflife", "stfeco", "stfgov", "stfedu","stfdem","","","","",rows = 6)
lay

graph_sem(model = fit.mimic.med,   
                    layout = lay,  
                    text_size = 1.5 ) 
```




# multi-group SEM EX3
The goal of such an analysis is to assess whether the relationships among predictor and response variables vary by group

1. compare the difference of models with unconstrained and constrained loadings 