---
title: "SEM_lab2"
output: pdf_document
---

```{r setup, include=FALSE}
### DATA MANIPULATION ###
library("haven")        
library("dplyr")      
library("psych")
library('stringr')

### MODELING ###
library("lavaan")       

### VISUALIZATION ###
library("tidySEM")
ess_df <- haven::read_sav("https://github.com/albertostefanelli/SEM_labs/raw/master/data/ESS4e04_5.sav")
ess_df
```

# EX1

```{r}
# 1.Fit the a measurement model with both welfare support and egalitarianism and review the fit statistics. Also review the parameter estimates. How close are they to the population values?
model_ws_eg <- '
## Welfare Support Factor ##
welf_supp =~ gvslvol+gvslvue+gvhlthc

## Egalitarianism ##
egual =~ gincdif+dfincac+smdfslv 
'

fit_ws_eg <- cfa(model_ws_eg, # model formula
                 data=ess_df)  # data frame
summary(fit_ws_eg)
```

1.there are 13 parameters, and total obs is 1760. 

2.from the test stats: chi-square is 40.125 with 8 degree of freedom, and p-value is very significant. 

3. from the latent variables: gvslvue is lower than others, and dfincac has negative lambda value. 

4. covariance of welf_supp and egual is negative (-0.225)




```{r}
# 2.Predict welfare support by adding the gender covariate and assess the changes in fit statistics and parameter estimates.
model_ws_eg1 <- '
## Welfare Support Factor ##
welf_supp =~ gvslvol+gvslvue+gvhlthc
welf_supp ~ gndr

## Egalitarianism ##
egual =~ gincdif+dfincac+smdfslv 
'

fit_ws_eg1 <- cfa(model_ws_eg1, # model formula
                 data=ess_df)  # data frame

summary(fit_ws_eg1)
```

```{r}
model_ws_eg2 <- '
## Welfare Support Factor ##
welf_supp =~ gvslvol+gvslvue+gvhlthc
welf_supp ~ agea + hinctnta + dfincac + gndr + eduyrs

## Egalitarianism ##
egual =~ gincdif+dfincac+smdfslv 
'

fit_ws_eg2 <- cfa(model_ws_eg2, # model formula
                 data=ess_df)  # data frame


model_ws_eg3 <- '
## Welfare Support Factor ##
welf_supp =~ gvslvol+gvslvue+gvhlthc
welf_supp ~ agea +dfincac + gndr + eduyrs

## Egalitarianism ##
egual =~ gincdif+dfincac+smdfslv 
'

fit_ws_eg3 <- cfa(model_ws_eg3, # model formula
                 data=ess_df)  # data frame

model_ws_eg4 <- '
## Welfare Support Factor ##
welf_supp =~ gvslvol+gvslvue+gvhlthc
welf_supp ~ agea  + gndr + eduyrs

## Egalitarianism ##
egual =~ gincdif+dfincac+smdfslv 
'
# 9 obs variable, 4 loadings for welf_supp&egual (since we restrict laoding=1 for the first variable of each latent variable), latent_var=2, covariance_latent=False, exogenous=3 (agea, gndr, eduyrs), then we can use these info to calculate the degree of freedom

fit_ws_eg4 <- cfa(model_ws_eg4, # model formula
                 data=ess_df)  # data frame

# model1:gndr
fitm_model_ws_eg1 <- fitMeasures(fit_ws_eg1, c("logl","AIC",  "BIC", "chisq","df","pvalue","cfi","tli", "rmsea"),  output = "matrix")

# model4: agea  + gndr + eduyrs
fitm_model_ws_eg4 <- fitMeasures(fit_ws_eg4, c("logl","AIC",  "BIC", "chisq","df","pvalue","cfi","tli", "rmsea"),  output = "matrix")

# model3:agea +dfincac + gndr + eduyrs
fitm_model_ws_eg3 <- fitMeasures(fit_ws_eg3, c("logl","AIC",  "BIC", "chisq","df","pvalue","cfi","tli", "rmsea"),  output = "matrix")

# model2: agea + hinctnta + dfincac + gndr + eduyrs
fitm_model_ws_eg2 <- fitMeasures(fit_ws_eg2, c("logl","AIC",  "BIC", "chisq","df","pvalue","cfi","tli", "rmsea"),  output = "matrix")

fitm_model_ws_eg1
fitm_model_ws_eg4
fitm_model_ws_eg3
fitm_model_ws_eg2


```
# comparing model with adding different covariances variable

model1: gender,  aic=33303.601, CFI(0.937), TLI(0.906) and RMSEA(0.064) are good, meet the criterion.

model4: agea  + gndr + eduyrs,aic=33261.924, CFI(0.889), TLI(0.848), RMSEA(0.067) lower than model1, and below the criterion.

model3:agea +dfincac + gndr + eduyrs, aic=33254.005, CFI(0.895), TLI(0.849), RMSEA(0.067) lower than model1.

model2: agea + hinctnta + dfincac + gndr + eduyrs, aic=29690.630, CFI(0.885), TLI(0.840), RMSEA(0.065) are lower than model1

model2, model3, model4 all have lower aic than model1, CFI, TLI and RMSEA are similar to each other, but does not meet the cut-off criterion. Model2 has lowest aic, can be consider to use compare to other models.

```{r}
degres_of_freedom <- function(obs_var=obs_var, # n observed variables 
loads=loads,     # n estimated factors loadings
regression=regression, # n regression
latent_var=latent_var, # n latent variables
covariance_latent=TRUE, # covariance latent vars
exogenous=exogenous){ # n exogenous 
# next function returns n combination of 2 or more variables 
# we use it to calculate the n covariance of the exogenous variables
comb = function(n, x) {
  if(n==0){return(0)} # if we have no exogenous variables return 0
  if(n==1){return(0)} # if we have 1 exogenous variable return 0
  else{
  factorial(n) / factorial(n-x) / factorial(x)}
}

# 1. pieces of information (observed variables)
info <- obs_var*(obs_var + 1)/2 

#2. measurement part = n factor loadings + n residual variances. 
measurement <- (loads) + obs_var

#3. structural part = n regressions + n residual factor variance.
structural <- ifelse(regression>0,0,regression) + latent_var

# 4. exogenous variables = n variances + n covariances.
# first, let's calculate the covariances 
cov <- NA
# lavaan removes by default the LV covariance when you include a regression
# this code takes into account lavaan behaviour
if(covariance_latent==TRUE){ 
cov<-comb(latent_var, 2)}else{cov<-0}

# second, let's get the n variances and n covariances for our exogenous vars.
exogenous_var <- exogenous + comb(exogenous, 2)

# finally, let's calculate the df
df = info - (measurement + structural + exogenous_var+ cov)
return(df)
} 

# for each model we have to find the number of obs, number of loadings, number of regression, number of latent variable, if we have covariance between latent variable, and exogenous variable
df_model1 <- degres_of_freedom(obs_var=6,loads=4,regression=0,latent_var=2,covariance_latent=TRUE,exogenous=0)


```

# EX2

```{r}
model_mediation <- '
## Welfare Support Factor ##
welf_supp =~ gvslvol+gvslvue+gvhlthc

## Egalitarianism ##
egual =~gincdif+dfincac+smdfslv 

## Direct effect(s) ##
welf_supp ~ c1*agea+c2*eduyrs+c3*gndr+c4*hinctnta

## Mediator ##
# Path A
egual ~ a1*agea
egual ~ a2*eduyrs
egual ~ a3*gndr
egual ~ a4*hinctnta

# Path B
welf_supp ~ b1*egual

## Indirect effect (a*b) ##
ab_age := a1*b1
ab_edu:=a2*b1
ab_gndr:=a3*b1
ab_hinc:=a4*b1

## Total effect ##
total_age:=c1+(a1*b1)
total_edu:=c2+(a2*b1)
total_gndr:=c3+(a3*b1)
total_hinc:=c4+(a4*b1)
'

fit_mediation=cfa(model_mediation, data=ess_df)
fitm_mediation_model <- fitMeasures(fit_mediation, c("logl","AIC",  "BIC", "chisq","df","pvalue","cfi","tli", "rmsea"),  output = "matrix")
fitm_mediation_model


# R^2
round(inspect(fit_mediation,"r2"),3)
# R^2 is the proportion of latent response explained by predictor


# how to calculate the direct and indirect effect manually
## Coefficients ## 
a <- inspect(fit_mediation,"est")$beta["egual","agea"]
b <- inspect(fit_mediation,"est")$beta["welf_supp","egual"]
## Standard Errors ## 
sigma_a <- inspect(fit_mediation,"se")$beta["egual","agea"]
sigma_b <- inspect(fit_mediation,"se")$beta["welf_supp","egual"]
## Manual calculation of indirect effect a x b ## 
a * b

## Manual calculation of the sigma a x b ## 
sqrt(a^2*sigma_b^2 + b^2*sigma_a^2)

## Manual calculation of the total effect (a x b) + c' ## 
a * b + inspect(fit_mediation,"est")$beta["welf_supp","agea"]


# calculate the direct and indirect effect by lavaan
tidy_results <- table_results(fit_mediation,      columns = c("label", "est_sig", "se"),digits = 2,)

tidy_results %>% filter(str_detect(label,"ab_age|total_age")) 
```

