---
title: "R Notebook"
output: html_notebook
---

```{r, include=FALSE}
### DATA MANIPULATION ###
library("haven")        
library("dplyr")      
library("psych")
library('stringr')

### MODELING ###
library("lavaan")       

### VISUALIZATION ###
library("corrplot")     
library("tidySEM")

```

# Data exploration

```{r}
ess_df <- haven::read_sav("https://github.com/albertostefanelli/SEM_labs/raw/master/data/ESS4_belgium.sav")
ess=ess_df%>%select(
  ## Welfare support items ##
                  gvslvol, # the old
                  gvslvue, # the unemployed
                  gvhlthc, # the sick
                  gvcldcr, # working parents
                  gvjbevn, # job for everyone
                  gvpdlwk, # paid sick leave  
                  ##    Economic criticism items ##
                  sbstrec, # strain on economy
                  sbbsntx, # too much taxes
                  ##    Social criticism items ## 
                  sbprvpv, # poverty
                  sbeqsoc, # more equal society
                  sbcwkfm, # work and family
                  ##    Moral criticism items ##
                  sblazy,  # people lazy 
                  sblwcoa, # care for others
                  sblwlka  # look after others
)
descriptive_ess <-as.data.frame(psych::describe(ess))

descriptive_ess <- dplyr::select(descriptive_ess, 
  n,mean,sd,median,min,max,skew,kurtosis)
descriptive_ess
```

# Welfare group

```{r}
ess_welfare=ess_df%>%select(
   ## Welfare support items ##
                  gvslvol, # the old
                  gvslvue, # the unemployed
                  gvhlthc, # the sick
                  gvcldcr, # working parents
                  gvjbevn, # job for everyone
                  gvpdlwk  # paid sick leave  
)
welfare_cov=cov(ess_welfare, use="pairwise.complete.obs")
welfare_cov

welfare_cor=cov2cor(welfare_cov)
welfare_cor

corrplot::corrplot(welfare_cor, is.corr=F, method="circle", type="upper", addCoef.col="black")
```

# 1 factor CFA model ( 3 indicators )

## Sorts of identification:

non-identified: df<0, impossible to estimate parameters 

over-identified: df>0, we should strive for this

just-identified: df=0, it is ok, but the fit cannot be assessed

empirical under-identification: it can happen when two or more indicators are highly correlated (multicollinearity)

## Degrees of freedom = Number of parameters to estimate - Pieces of information
## Pieces of information = p(p+1)2=3(3+1)2=6.

where p is the number of “manifest” indicators i (also called “measured variables” or “observed indicators”), in this case gvslvol gvslvue, gvhlthc.

## Number of parameters to estimate = ψf+λji+θi=1+2+3=6
where ψf is the latent factor f variance, λfi is the loading on the latent factor f for each observed indicator i, and θi is the residual variance for each observed indicator i.

## Degrees of freedom = Number of parameters to estimate - Pieces of informations =6−6=0

# CFA model
```{r}
model1='wel_supp =~ gvslvol + gvslvue + gvhlthc'
fit.model1=cfa(model1, data=ess)
summary(fit.model1)

#fix one factor loading to 1
#fix factor variance to 1, freely estimate the first factor loadings
#fix the average loading to 1
model2='welf_supp =~ NA*gvslvol + gvslvue + gvhlthc
welf_supp~~1*welf_supp'
fit.model2=cfa(model2, data=ess)
```

## different between these 2 models

1. variances of wel_supp changes
2. the latent variables have different loadings

# 1 factor model with 6 indicators
```{r}
model3 <-'welf_supp =~ gvslvol + gvslvue + gvhlthc + gvcldcr + gvjbevn + gvpdlwk'


fit.model3 <- cfa(model3,# modelformula
                data = ess) # data frame 
```

## why does the degree of freedom changes from 6 to 9?

# Model Diagnostics

1.Estimates
Do the indicators load well on the factor(s)?

2.Model fit
Are the fit indices good?

3.Heywood cases/ reasonable solution
Are variances positive?
Are r2 is below 1?
Are there any extreme standard errors?

```{r}
# observed covariance matrix
lavInspect(fit.model3, "sampstat")

# estimated covariance matrix
fitted(fit.model3)

# difference between the observed and estimated covariance matrix
residuals(fit.model3)

# standard errors for the free parameters in the model3
lavTech(fit.model3, "se")

# fit statistics for model3
fitMeasures(fit.model3)

# modification indices
modificationIndices(fit.model3)

# parameters estimates with confidence intervals
parameterEstimates(fit.model3, standardized=TRUE)

# using fitted model3, return commonly used fit measures, and indicates the standardized result
summary(fit.model3, fit.measures=T, standardized=T)

# extract the standrardized factor loading
inspect(fit.model3, what='std')$lambda

#more info on factor loading, extract all the estimated parameters
tidy_results <- table_results(fit_ws_6,             
  columns = c("label", "est_sig", "se", "confint"),
  digits = 2)

tidy_results %>% filter(str_detect(label,"welf_supp."))

#residual variances (theta)
theta <- round(inspect(fit.model3, "est")$theta,3)
theta.std <- round(inspect(fit.model3, "std")$theta,3) 
r2 <- round(inspect(fit.model3, "r2"),3)

data.frame(row.names = c(),# empty the columns names
           Variables = colnames(theta), # variable names 
           "Residuals" = diag(theta),# diagonal theta
           "Std. Residuals" = diag(theta.std),# diagonal std. theta
           "R Squared" = r2) # R-squared
```
# Model fit statistics

Global fit measures:
1.They take into account how the entire entire model fit the data

2.(some) rules of thumb: CFI/TLI>0.95,RMSEA<0.05,SRMR<0.06

current practice is: chi-square value + df + pvalue, RMSEA, CFI and SRMR

we can modify the model based on a review of:
MI’s in combination with EPC’s (Expected Value Change). Both need to be “substantial”

Theory or the source of the data

```{r}
#global fit measures
fitMeasures(fit.model3, c("logl","AIC", "BIC", "chisq", "df", "pvalue", "cfi", "tli","rmsea"), output = "matrix")

#modification indices
mi <- inspect(fit.model3,"mi")
mi.sorted <- mi[order(-mi$mi),] # sort from high to low mi.sorted[1:5,] # only display some large MI values
mi.sorted[1:5,] # only display some large MI values

#plot modification indices
plot(mi.sorted$mi)
abline(h=3.84) #add horizontal reference line (chi-square value for 1 df and p=0.05)
```

# 3-factor CFA model

```{r}
ess_df_selected <- ess_df %>% select(
                  ## Economic criticism items ##
                  sbstrec, # strain on economy
                  sbbsntx, # too much taxes
                  ##    Social criticism items ## 
                  sbprvpv, # poverty
                  sbeqsoc, # more equal society
                  sbcwkfm, # work and family
                  ##    Moral criticism items ##
                  sblazy,  # people lazy 
                  sblwcoa, # care for others
                  sblwlka  # look after others
)

welfare_crit_cov <- cov(ess_df_selected, use = "pairwise.complete.obs")

welfare_crit_cor <- cov2cor(welfare_crit_cov)

corrplot::corrplot(welfare_crit_cor, 
                   is.corr = FALSE,       # whether is a correlation matrix 
                   method = "circle",     # magnitude of covariance as circles 
                   type = "upper",        # remove the bottom of the covariance matrix
                   addCoef.col = "black"  # add to the plot the coefficients
          )

#build the model
model_wc <-'
## Economic criticism ##
wc_econo =~ sbstrec + sbbsntx
## Social criticism ## 
wc_socia =~ sbprvpv + sbeqsoc + sbcwkfm
##  Moral criticism ##
wc_moral =~ sblazy + sblwcoa + sblwlka
'

fit_wc <- cfa(model_wc,              # model formula
             data = ess_df_selected  # data frame
  )
summary(fit_wc, standardized=TRUE)

# plot the sem model
lay <- get_layout("wc_econo", "", "", "wc_socia","", "","wc_moral", "",
                  "sbstrec", "sbbsntx", "sbprvpv", "sbeqsoc", "sbcwkfm", "sblazy", "sblwcoa", "sblwlka", rows = 2)

#method 1 to plot
plot_wc <- graph_sem(model = fit_wc,      # model fit
                    layout = lay,         # layout 
                    angle = 170           # adjust the arrows 
                    #label = "est_std",  # get standardized results (not rounded)
          )   
plot_wc


# method 2 to plot, For instance, it is possible to highlighting a specific model element, such as the low factor loading for sbcwkfm on wc_socia
graph_data <- prepare_graph(fit_wc)
 
edges(graph_data) <- graph_data %>% 
  edges() %>%
  mutate(colour = "black") %>%
  mutate(colour = replace(colour, from == "wc_socia" & to == "sbcwkfm", "red"))

plot(graph_data,
     layout = lay,        # layout 
     #label = "est_std",   # get standardized results (not rounded)
     angle = 170          # adjust the arrows 
  )
```

